
## ReDos

The regular expression denial of service ([ReDoS](https://en.wikipedia.org/wiki/ReDoS)) is an algorithmic complexity attack that produces a denial-of-service by providing a regular expression that takes a very long time to evaluate.

### StackOverflow.com Outage - July 20, 2016

The [blog post](http://stackstatus.net/post/147710624694/outage-postmortem-july-20-2016) explained the direct cause of the outage. An regular expressions, which is used to trim unicode space from start and end of a line, consume high CPU on the web servers.

```{r}

pattern = enc2utf8("^[\\s\u200c]+|[\\s\u200c]+$")

times.list <- list()
index = 1
seq_ = seq(1000,9000,by = 1000)
for(N in seq_){
    # string and pattern
    string = paste0("-- play happy sound for player to enjoy. ", paste0(rep(" ",N), collapse = ""),"boom!")
    
    # pre-compiled RE2
    regexp = re2(pattern)
    
    # run gc
    invisible(gc())
    
    # benchmark
    N.times <- microbenchmark(
        ICU = stri_detect_regex(string, pattern = pattern),
        PCRE = grepl(pattern, string, perl = TRUE),
        TRE = grepl(pattern, string, perl = FALSE),
        RE2n = re2_detect(string, pattern),
        RE2c = re2_detect(string, regexp),
        times = 4)
    
    times.list[[index]] <- data.frame(N, N.times)
    index= index+1
}

times <- do.call(rbind, times.list)

direct.label(linear.legend(times, limit=c(0,9000), breaks = seq_,xlab_name = "string size", title = "StackOverflow.com Outage"), "last.polygons")


```

### Java Classname

```{r}

pattern = "^(([a-z])+.)+[A-Z]([a-z])+$"

times.list <- list()
index = 1
seq_ = seq(1,24,by = 1)
for(N in seq_){
    # string and pattern
    string = paste0(paste0(rep("a",N), collapse = ""),"boom!")
    
    # pre-compiled RE2
    regexp = re2(pattern)
    
    # run gc
    invisible(gc())
    
    # benchmark
    N.times <- microbenchmark(
        ICU = stri_detect_regex(string, pattern = pattern),
        PCRE = grepl(pattern, string, perl = TRUE),
        TRE = grepl(pattern, string, perl = FALSE),
        RE2n = re2_detect(string, pattern),
        RE2c = re2_detect(string, regexp),
        times = 4)
    
    times.list[[index]] <- data.frame(N, N.times)
    index= index+1
}

times <- do.call(rbind, times.list)

direct.label(linear.legend(times, limit=c(0,24), breaks = seq_,xlab_name = "string size", title = "Java Classname Malicious Regex"), "last.polygons")

```

### "a?a"

Derived from Toby Dylan Hocking [tdhock/regex-tutorial
](https://github.com/tdhock/regex-tutorial)

```{r, echo=TRUE}

times.list <- list()

for(N in c(1:25)){
    # string and pattern
    string <- paste(rep("a", N), collapse="")
    pattern <- paste(rep(c("a?", "a"), each=N), collapse="")
    
    # pre-compiled RE2
    regexp = re2(pattern)
    
    # run gc
    invisible(gc())
    
    # benchmark
    N.times <- microbenchmark(
        ICU = stri_detect_regex(string, pattern = pattern),
        PCRE = grepl(pattern, string, perl = TRUE),
        TRE = grepl(pattern, string, perl = FALSE),
        RE2n = re2_detect(string, pattern),
        RE2c = re2_detect(string, regexp),
        times = 5)
    
    times.list[[N]] <- data.frame(N, N.times)
}

times <- do.call(rbind, times.list)

direct.label(linear.legend(times, xlab_name = "Pattern size", title = "pattern: 'a?a?aa' 'a?a?a?aaa' ..."), "last.polygons")
direct.label(log.legend(times, xlab_name = "pattern size", title = "Pattern: 'a?a?aa' 'a?a?a?aaa' ... log scale"), "last.polygons")

```

```{r}
string <- paste(rep("a", 27), collapse="")
pattern <- paste(rep(c("a?", "a"), each=27), collapse="")

system.time(re2_detect(string,pattern))
system.time(stri_detect_regex(string,pattern))
```

