
## Fail to Match

These strings will fail to match the pattern, and they test the speed of failed match.

```{r}
gen_bench = function(range_, pattern) {
    times.list <- list()
    index = 1
    for (N in range_) {
        # pre-compiled RE2
        regexp = re2(pattern)
        string = stringi::stri_rand_strings(1, N,
                                            pattern = "[a-zA-Z0-9 !@#$%()*Ω≈ç√∫˜µåœ∑†¨ˆ˙©ƒ]")
        
        # run gc
        invisible(gc())
        
        # benchmark
        N.times <- microbenchmark(
            ICU = stri_detect_regex(string, pattern = pattern),
            PCRE = grepl(pattern, string, perl = TRUE),
            TRE = grepl(pattern, string, perl = FALSE),
            RE2n = re2_detect(string, pattern),
            RE2c = re2_detect(string, regexp),
            times = 5
        )
        
        times.list[[index]] <- data.frame(N, N.times)
        index = index + 1
    }
    
    times <- do.call(rbind, times.list)
    return(times)
}

plot_bench = function(pattern) {
    res = gen_bench(seq(from = 10000, to = 70000, by = 10000), pattern)
    
    direct.label(
        linear.legend(
            res,
            limit = c(0, 71000),
            breaks = seq(from = 10000, to = 70000, by = 10000),
            xlab_name = "string size"
        ),
        "last.polygons"
    )
}
```

These two are easy because they start with an A,
giving the search loop in something to call `memchr` in C++. These tests are from RE2 library.

```{r}
EASY0 = "ABCDEFGHIJKLMNOPQRSTUVWXYZ$"
plot_bench(EASY0)

EASY1 = "A[AB]B[BC]C[CD]D[DE]E[EF]F[FG]G[GH]H[HI]I[IJ]J$"
plot_bench(EASY1)
```

This is a little harder, since it starts with a character class
and thus can't be `memchr`'ed in C++.

```{r}
MEDIUM = "[XYZ]ABCDEFGHIJKLMNOPQRSTUVWXYZ$"
plot_bench(MEDIUM)
```

This is a fair amount harder, because of the leading [ -~]*.
A bad backtracking implementation will take O(text^2) time to
figure out there's no match.

```{r}
HARD = "[ -~]*ABCDEFGHIJKLMNOPQRSTUVWXYZ$"
plot_bench(HARD)
```

This stresses engines that are trying to track parentheses.

```{r}
PARENS = "([ -~])*(A)(B)(C)(D)(E)(F)(G)(H)(I)(J)(K)(L)(M)(N)(O)(P)(Q)(R)(S)(T)(U)(V)(W)(X)(Y)(Z)$"
plot_bench(PARENS)
```

This pattern has a high degree of fanout.
NFA execution will be particularly slow.

```{r}
FANOUT = "(?:[\\x{80}-\\x{10FFFF}]?){100}[\\x{80}-\\x{10FFFF}]"
range_ = seq(from = 10000, to = 80000, by = 10000)
times.list <- list()
index = 1
for(N in range_){
    # pre-compiled RE2
    regexp = re2(FANOUT)
    pattern = FANOUT
    string = stringi::stri_rand_strings(1,N, 
                                        pattern="[a-zA-Z0-9 !@#$%()*Ω≈ç√∫˜µåœ∑†¨ˆ˙©ƒ]")
    
    # run gc
    invisible(gc())
    
    # benchmark
    N.times <- microbenchmark(
        ICU = stri_detect_regex(string, pattern = pattern),
        PCRE = grepl(pattern, string, perl = TRUE),
        RE2n = re2_detect(string, pattern),
        RE2c = re2_detect(string, regexp),
        times = 5)
    
    times.list[[index]] <- data.frame(N, N.times)
    index = index+1
}

times <- do.call(rbind, times.list)

direct.label(linear.legend(times, limit = c(0,90000), breaks = range_, xlab_name = "string size"), "last.polygons")
```
